---
title: "Conditional Power Promising Zone Sample Size Re-estimation Inflates Type I Error in Single-Arm Binary Trials"
subtitle: "An Exact-Enumeration Study and Comparison with Bayesian Predictive Probability SSR"
author: "Author(s) anonymized for peer review"
date: today
abstract: |
  The Mehta-Pocock conditional-power (CP) promising-zone framework
  is widely used for sample size re-estimation, with Type I error
  control derived for two-arm trials with continuous test statistics
  using an inverse-normal combination test. We examine a common
  extrapolation of this framework to single-arm binary trials using
  a one-sided one-proportion z-test final analysis. Exact enumeration
  of the joint interim-and-final binomial distribution shows that
  this extrapolation can inflate Type I error. In a representative
  oncology setting, exact Type I error exceeded the nominal 0.05
  level for every promising-zone lower threshold evaluated. Across
  five design configurations, the z-test critical count's
  discreteness bias changed sign, so thresholds controlling Type I
  error in one setting did not necessarily control it in another.
  In contrast, Bayesian predictive-probability sample size
  re-estimation with separate interim and final posterior thresholds
  produced a monotone, directly calibratable Type I error curve.
  A symmetric comparison showed that pairing CP-driven sample size
  re-estimation with a Bayesian posterior final analysis restored
  calibration, identifying the final decision rule as a sufficient
  cause of the observed inflation. These results support
  design-specific exact Type I error reporting for single-arm binary
  promising-zone designs, or replacement of the z-test final analysis
  with a calibrated alternative.
keywords: [adaptive design, sample size re-estimation, conditional power, predictive probability, single-arm phase II, Type I error]
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    self-contained: true
    code-fold: true
    code-summary: "Show code"
    fig-width: 9
    fig-height: 5.5
  pdf:
    documentclass: scrartcl
    # No titlepage option: keep the abstract on the cover page with the
    # title/author/date. The \tableofcontents redefinition below forces a
    # page break after the abstract so the TOC starts on its own page.
    papersize: letter
    geometry:
      - margin=1in
    fontsize: 11pt
    linestretch: 1.25
    colorlinks: true
    linkcolor: NavyBlue
    citecolor: NavyBlue
    urlcolor: NavyBlue
    toc: true
    toc-depth: 2
    number-sections: true
    # Match the HTML rendering: same width/aspect ratio + PNG (rasterized,
    # anti-aliased) device so plots in the PDF look the same as in the HTML.
    fig-width: 9
    fig-height: 5.5
    fig-format: png
    fig-dpi: 200
    fig-pos: "H"
    fig-cap-location: bottom
    tbl-cap-location: top
    code-overflow: wrap
    keep-tex: true
    echo: false
    include-in-header:
      text: |
        \usepackage{float}
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{array}
        \usepackage{pdflscape}
        % Force a clean page break after the abstract so that the TOC
        % starts on its own page (page 2). This keeps the title block
        % and abstract together on the cover page.
        \let\OriginalTOC\tableofcontents
        \renewcommand{\tableofcontents}{\clearpage\OriginalTOC}
        % Uncomment for journal-review version with line numbers:
        % \usepackage{lineno}
        % \linenumbers
execute:
  echo: true
  warning: false
  message: false
  cache: true
bibliography: references.bib
---

```{r setup}
#| include: false
library(tidyverse)
library(gt)
library(scales)
library(patchwork)   # plot composition
library(glue)        # string interpolation
library(latex2exp)   # TeX() for math in ggplot subtitles/labels

# All helper, simulator, and exact-OC functions are defined in R/analysis.R.
# They are sourced here so the manuscript code chunks below stay focused on
# parameter setup, sweeps, and presentation. We resolve the script's path
# relative to the qmd file so the render works regardless of the working
# directory at the time `quarto render` is invoked.
.analysis_path <- local({
  candidates <- c(
    "R/analysis.R",                       # running from the repo root
    "cp-ssr-binary-trials/R/analysis.R",  # running from one level above
    file.path(dirname(knitr::current_input(dir = TRUE)), "R", "analysis.R")
  )
  hit <- candidates[file.exists(candidates)]
  if (length(hit) == 0L) {
    stop("Could not locate R/analysis.R. Run `quarto render` from the ",
         "CP-SSR-Binary-Trials repo root, or place R/analysis.R next to ",
         "the .qmd file.")
  }
  hit[[1L]]
})
source(.analysis_path)

set.seed(20260414)
# N_SIM = 100,000 is the outline target. Use 5,000–50,000 for development
# (CP and Bayesian sweeps + generalizability scale roughly linearly in N_SIM).
N_SIM <- 10000L
```

{{< pagebreak >}}

# Introduction

Single-arm Phase II oncology trials with a binary objective response
rate (ORR) endpoint are a long-standing design framework for assessing
biological activity of investigational agents [@simon1989]. In this
setting, the trial sponsor must commit to a sample
size before observing data on the new agent, and yet the assumed effect
size, and therefore the planned power, is rarely well-calibrated. Sample
size re-estimation (SSR) addresses this by allowing the planned $N$ to be
adjusted at an interim analysis based on observed information.

Among adaptive SSR strategies, the **conditional-power promising zone**
framework introduced by @mehta2011 is the most widely used in industry
practice. Its appeal is partly statistical and partly regulatory: in the
two-arm trial with a normally distributed test statistic, the
inverse-normal combination test of @lehmacher1999 and @muller2001 provides
an *analytic* guarantee that the Type I error is preserved when sample
size is re-estimated based on a continuous interim test statistic
[@mehta2011]. The 2019 FDA adaptive design guidance [@fda2019]
explicitly recognises this design family and requires demonstration of
Type I error control, analytically where possible, otherwise via
simulation.

A practical question, however, is whether the analytic guarantee carries
over to single-arm trials with binary endpoints. The interim test statistic
is no longer a continuous $z$ statistic but a binomial response count,
which lives on a coarse discrete grid. The combination-test machinery that
underpins the two-arm proof has no obvious analogue in this setting. To
date, this gap appears to have been bridged in practice by importing the
promising-zone *form* of the design and verifying the SSR through
parameter tuning, on the assumption that what works for normal endpoints
also works for binary ones.

In this paper we show, by exact enumeration, that this assumption
fails for the natural CP-promising-zone implementation studied here.
We use a one-sided one-proportion $z$-test (normal approximation) for
the final analysis and the standard Mehta–Pocock SSR target rule
(extend $N$ until CP rises back to $\text{CP}_U$); these choices
matter for the magnitudes of the inflation we report and we discuss
their role in Section 5. For single-arm binary trials calibrated to a
representative oncology setting ($p_0 = 0.23$, $p_1 = 0.35$,
one-sided $\alpha = 0.05$, $N_\text{init} = 84$, $n_\text{int} = 65$,
$N_\text{max} = 200$), exact Type I error sits **uniformly above**
$\alpha = 0.05$ across the entire promising-lower-bound grid we
examine, varying only mildly with $\text{CP}_L$ (every value below
is computed exactly; the variation is the genuine effect of moving
$\text{CP}_L$, not sampling error). No value of $\text{CP}_L$ on the
grid achieves nominal Type I error control in this setting. We provide a mechanistic
explanation: at finite $N$, the one-sided $z$-test critical count is
the smallest *integer* with $z \geq z_\alpha$, so the actual Type I
error of the underlying fixed-$N$ test differs from the nominal
$\alpha$ by a *signed* amount whose sign and magnitude depend on
$(p_0, N)$, it can be either upward (e.g., the base setting) or
downward (e.g., the $p_0 = 0.10$, $N_\text{init} = 55$ setting in
@fig-generalizability). SSR moves trials between such finite-$N$ tests, with their
own opposite-sign discreteness biases, in a way that depends on
$\text{CP}_L$ and the realised interim count. The resulting T1E
behaviour is design-specific and cannot be controlled by a
fixed-$\text{CP}_L$ rule that generalises across designs.

We contrast this behaviour with Bayesian predictive-probability (PP) SSR
[@dmitrienko2006; @lee2008; @saville2014; @liu2023], which operates on
a near-continuous posterior probability and exhibits monotone exact
Type I error as a function of the final-analysis posterior threshold
$\gamma_\text{final}$ (the curve is still stepwise on the integer
response-count grid, but the surface is far less irregular than the
CP_L surface and admits direct calibration). The Bayesian PP design can be
calibrated to any target Type I error directly from the exact operating
characteristics, with separate interim and final thresholds
($\gamma_\text{int}$, $\gamma_\text{final}$) governing distinct
decision points.

Our contribution is threefold: (1) we document, by exact computation
and simulation, that CP promising-zone SSR cannot be calibrated to
nominal $\alpha$ by a fixed-$\text{CP}_L$ rule that generalises across
single-arm binary designs, in the base setting no $\text{CP}_L$ on
the grid achieves $\text{T1E} \leq \alpha$, and across a
generalizability grid the discreteness bias changes sign so that
$\text{CP}_L$ values that control T1E in one configuration may not
control it in another; (2) we explain the mechanism analytically in
terms of the discrete $z$-test critical count's design-specific
rounding bias combined with per-integer boundary crossings; and
(3) we show that Bayesian PP SSR, for which exact Type I error is
monotone in the final-analysis threshold $\gamma_\text{final}$ and
admits direct calibration on a far less irregular surface than the
CP_L grid, is more straightforward to calibrate in the settings
studied, and is a practically attractive alternative when calibrated
Type I error control is required. The remainder of the
paper is organised as follows. Section 2 reviews the relevant background.
Section 3 specifies the methods. Section 4 presents results.
Sections 5 and 6 discuss limitations and conclude.


# Background


## Single-Arm Phase II Design

Single-arm Phase II oncology trials test the response rate $p$ of a new
agent against a fixed null $p_0$, where $p_0$ is typically obtained from
historical-control data [@simon1989]. The classical frequentist design is
Simon's two-stage [@simon1989], which fixes both $N$ and a single interim
look but does not include sample-size re-estimation.

Bayesian alternatives operate on the posterior $P(p > p_0 \mid x, n)$
under a Beta-Binomial model. @lee2008 introduced the predictive
probability design, in which the trial monitors the posterior probability
that the trial will *eventually* declare success at the planned final
analysis. The predictive probability under a Beta($a, b$) prior and
binomial likelihood is available in closed form as a Beta-Binomial sum
[@lee2008; @saville2014].

The choice between conditional power (CP) and predictive probability (PP)
for interim monitoring has been debated since at least @spiegelhalter1986.
CP projects post-interim data under a fixed alternative $p_1$ and asks how
likely the trial is to reach significance. PP integrates over the
posterior on $p$ and asks how likely the trial is to reach significance
*given the data observed so far*. In single-arm binary trials, where the
posterior on $p$ is well-defined and computationally cheap, PP is a
natural quantity to monitor [@saville2014; @lan2009].


## Sample Size Re-estimation

@mehta2011 introduced the **conditional-power promising zone** framework
for SSR. At an interim analysis, CP is computed under the alternative
$p_1$. Three thresholds, futility ($\text{CP}_\text{fut}$), promising
lower ($\text{CP}_L$), and promising upper ($\text{CP}_U$), partition
the CP scale into futility, unfavorable, promising, and favorable zones.
In the promising zone the planned $N$ is increased to a value chosen so
that CP at the new $N$ rises to $\text{CP}_U$ (capped at a budget
$N_\text{max}$). The original Mehta–Pocock proof of Type I error control
relies on the inverse-normal combination test of @lehmacher1999 and
@muller2001, which combines pre- and post-interim test statistics with
fixed weights that are *invariant* to the SSR decision. This invariance
is what preserves the joint distribution of the combined statistic under
$H_0$ and yields analytic Type I error control.

Bayesian predictive monitoring of clinical trials was developed by
@dmitrienko2006 and later extended and applied in single-arm Phase II
and interim-monitoring contexts by @lee2008 and @saville2014;
@liu2023 subsequently studied predictive-power-based SSR specifically
for Phase 2 dose-finding. We adopt this lineage's PP design here. The PP design uses two independent thresholds: an *interim
efficacy* bar $\gamma_\text{int}$ (stop early for success if
$P(p > p_0 \mid x_\text{int}, n_\text{int}) \geq \gamma_\text{int}$)
and a *final analysis* bar $\gamma_\text{final}$ (declare success at
the end of the trial if $P(p > p_0 \mid x_\text{total}, n_\text{final})
\geq \gamma_\text{final}$). The predictive probability used as the
interim SSR trigger is computed *under* $\gamma_\text{final}$, not
$\gamma_\text{int}$. Because Type I error in this design is a continuous
function of $\gamma_\text{final}$, calibration is straightforward
either by simulation or, as we use throughout Section 4, by exact
enumeration.

The key gap motivating the present paper is that the inverse-normal
combination test that anchors the Mehta–Pocock proof has no obvious
analogue when the interim test statistic is a binomial response count
on a coarse discrete grid. Practitioners frequently transport the
promising-zone *form* of the CP design to the single-arm binary setting
by analogy, on the assumption that simulation-tuned zone boundaries can
recover Type I error control. We test that assumption directly.


# Methods


## Design Parameters

We consider a single-arm Phase II oncology trial with a binary objective
response rate (ORR) endpoint, comparing the response rate $p$ against a
historical null $p_0$,

$$H_0: p \leq p_0 \quad \text{vs.} \quad H_1: p = p_1 .$$

An interim analysis is conducted at $n_\text{int}$ patients. Depending
on the observed data, the trial either stops early (efficacy or futility)
or continues, with the option to extend enrollment up to a budget cap
$N_\text{max}$. The base-case design parameters reflect a representative
single-arm Phase II oncology setting and are summarised below; the same
parameter object is used by every computation in the rest of the paper.

```{r params}
# ── Fixed design parameters (representative oncology setting) ─────────────
p0      <- 0.23   # null response rate (historical control)
p1      <- 0.35   # target response rate
alpha   <- 0.05   # one-sided Type I error target
power   <- 0.80   # target power
n_init  <- 84L    # planned N (one-sided alpha = 0.05 hypothesis test)
n_int   <- 65L    # interim analysis (absolute N)
n_max   <- 200L   # hard budget cap

# ── Bayesian prior ────────────────────────────────────────────────────────
ap <- 0.5; bp <- 0.5   # Jeffreys Beta(0.5, 0.5)

# ── CP zone parameters, swept in Section 3 ───────────────────────────────
cp_upper    <- 0.80   # fixed throughout; favorable zone above this
cp_futility <- 0.10   # fixed; stop for futility below this

# ── Bayesian thresholds, swept in Section 4 ─────────────────────────────
gamma_int   <- 0.99   # interim efficacy stop bar, kept high so the interim
                      # branch contributes only a small share of T1E. At
                      # n_int = 65, P(post >= 0.95 | p = p0) is already ~5.5%,
                      # which by itself would consume more than alpha; 0.95
                      # is too lenient once the interim stop is active, so we
                      # set it to 0.99 (consistent with the outline).

# ── Scenario grid: null + insurance region + alternative ─────────────────
delta     <- p1 - p0
scenarios <- round(c(
  p0,
  p0 + 0.25 * delta,
  p0 + 0.50 * delta,
  p0 + 0.75 * delta,
  p1,
  p1 + 0.25 * delta
), 4)

scen_labels <- c(
  "p₀",
  "p₀ + 0.25Δ",
  "p₀ + 0.50Δ",
  "p₀ + 0.75Δ",
  "p₁",
  "p₁ + 0.25Δ"
)

# Lookup table for joining labels, avoids fragile rep() calls throughout
scen_label_tbl <- tibble(p_true = scenarios, Label = scen_labels)

# ── CP promising lower sweep ──────────────────────────────────────────────
cp_lower_grid <- seq(0.30, 0.80, by = 0.05)

# ── Bayesian gamma_final sweep, narrowed to the calibration-relevant range ──
gamma_fin_grid <- seq(0.95, 0.99, by = 0.005)
```


## Helper Functions

All computations rely on the following helpers. `x_crit_z_fn` returns the
minimum integer response count at which the one-sided one-proportion
$z$-test rejects $H_0$ at level $\alpha$, this is the same critical
count used for the final analysis in the CP design. `cp_fn` projects
post-interim data under $p_1$ and computes conditional power against
that critical count. `ssr_cp_fn` implements the Mehta–Pocock SSR rule:
in the promising zone, search upward from $N_\text{init}$ for the
smallest $n$ such that CP at $n$ rises back to $\text{CP}_U$, capping
at $N_\text{max}$. `pred_prob_fn` and `post_prob` compute the
Beta-Binomial predictive and posterior probabilities used by the
Bayesian PP design under prior Beta($a_p, b_p$) = Beta($0.5, 0.5$).

```{r helpers}
#| eval: false
# These functions are loaded from R/analysis.R via the setup chunk. They
# are reproduced here for inline transparency in the HTML version of the
# manuscript. Edit R/analysis.R if you want to change behavior.

# ── Critical response count: minimum x such that posterior >= thresh ──────
# (Bayesian: used by pred_prob_fn for the Bayesian final analysis)
x_crit_fn <- function(n, p0, ap, bp, thresh) {
  for (x in 0L:n) {
    if (1 - pbeta(p0, ap + x, bp + n - x) >= thresh) return(x)
  }
  return(n + 1L)
}

# ── Frequentist z-test critical count ─────────────────────────────────────
# Min integer x such that the one-sided one-proportion z-test rejects H0:p<=p0
# at level `alpha`, i.e., (x/n - p0)/sqrt(p0(1-p0)/n) > qnorm(1 - alpha).
# This MUST match the final-analysis decision rule used in sim_trial_cp().
x_crit_z_fn <- function(n, p0, alpha) {
  thresh <- n * p0 + qnorm(1 - alpha) * sqrt(n * p0 * (1 - p0))
  as.integer(floor(thresh) + 1L)
}

# ── Bayesian posterior probability P(p > p0 | x, n) ──────────────────────
post_prob <- function(x, n, p0, ap, bp) {
  1 - pbeta(p0, ap + x, bp + n - x)
}

# ── Beta-Binomial predictive probability of eventual success ──────────────
# Computed under gamma_fin (the FINAL analysis bar)
pred_prob_fn <- function(x_int, n_int, n_final, p0, ap, bp, gamma_fin) {
  a_post <- ap + x_int
  b_post <- bp + n_int - x_int
  n_rem  <- n_final - n_int
  xc     <- x_crit_fn(n_final, p0, ap, bp, gamma_fin)
  if (xc > n_final) return(0)
  x_need <- max(0L, xc - x_int)
  if (x_need > n_rem) return(0)
  sum(vapply(x_need:n_rem, function(xr) {
    exp(
      lchoose(n_rem, xr) +
      lbeta(a_post + xr, b_post + n_rem - xr) -
      lbeta(a_post, b_post)
    )
  }, numeric(1)))
}

# ── Conditional Power (frequentist) ──────────────────────────────────────
# Projects post-interim data under the alternative p1, using the same
# z-test critical count that the final analysis uses (no Bayesian boundary).
cp_fn <- function(x_int, n_int, n_final, p0, p1, alpha) {
  xc     <- x_crit_z_fn(n_final, p0, alpha)
  x_need <- max(0L, xc - x_int)
  n_rem  <- n_final - n_int
  if (x_need > n_rem) return(0)
  1 - pbinom(x_need - 1L, n_rem, p1)
}

# ── CP-based SSR (Mehta–Pocock promising zone rule) ──────────────────────
# Given that CP at n_init falls in the promising zone [cp_lower, cp_upper),
# search upward for the smallest n_final >= n_init such that CP(n_final)
# climbs back to cp_upper, capped at n_max.
ssr_cp_fn <- function(x_int, n_int, p0, p1, alpha,
                      cp_upper, n_init, n_max) {
  for (n_cand in seq(n_init, n_max)) {
    cp <- cp_fn(x_int, n_int, n_cand, p0, p1, alpha)
    if (cp >= cp_upper) return(n_cand)
  }
  return(n_max)
}

# ── Bayesian PP-based SSR: find min n_final such that PP >= pp_upper ──────
# pp_upper = 0.50 (favorable zone threshold)
ssr_bayes_fn <- function(x_int, n_int, p0, ap, bp, gamma_fin,
                         pp_upper, n_init, n_max) {
  for (n_cand in seq(n_init, n_max)) {
    pp <- pred_prob_fn(x_int, n_int, n_cand, p0, ap, bp, gamma_fin)
    if (pp >= pp_upper) return(n_cand)
  }
  return(n_max)
}
```


## CP Promising Zone SSR

At the interim ($n = n_\text{int}$) we compute conditional power under
the alternative $p_1$,

$$\text{CP}(x_\text{int}, n_\text{int}, N) =
P\!\left(X_\text{rem} \geq x_\text{crit}(N) - x_\text{int}
\;\big|\; X_\text{rem} \sim \text{Bin}(N - n_\text{int}, p_1)\right),$$

where $x_\text{crit}(N)$ is the $z$-test critical count at sample size
$N$ and the same one-sided $\alpha$ that the final analysis uses. The
trial then proceeds according to

$$
\text{Action} = \begin{cases}
\text{Futility stop} & \text{if } \text{CP} < \text{CP}_\text{fut} \\
\text{Unfavorable: continue to } N_\text{init} & \text{if } \text{CP}_\text{fut} \leq \text{CP} < \text{CP}_L \\
\text{Promising: SSR to } N' \leq N_\text{max} & \text{if } \text{CP}_L \leq \text{CP} < \text{CP}_U \\
\text{Favorable: continue to } N_\text{init} & \text{if } \text{CP} \geq \text{CP}_U
\end{cases}
$$

with $\text{CP}_U$ fixed at $0.80$ (the target power) and
$\text{CP}_\text{fut} = 0.10$ throughout. The promising-zone SSR rule
extends $N$ until CP rises back to $\text{CP}_U$, capped at
$N_\text{max}$. The final analysis is a one-sided one-proportion
$z$-test at $\alpha = 0.05$, using the same critical count that
appears in $\text{CP}_\text{fn}$ ensures interim and final decisions
operate on the same scale. We sweep the promising lower bound
$\text{CP}_L$ over the grid $\{0.30, 0.35, \ldots, 0.80\}$.

```{r cp-trial-sim}
#| eval: false
# Loaded from R/analysis.R; kept here as inline documentation.
sim_trial_cp <- function(p_true, cp_lower, cp_upper = 0.80,
                         cp_futility = 0.10, n_init = 84L,
                         n_int = 65L, n_max = 200L,
                         p0 = 0.23, p1 = 0.35, alpha = 0.05,
                         ap = 0.5, bp = 0.5) {
  x_int    <- rbinom(1L, n_int, p_true)
  cp_obs   <- cp_fn(x_int, n_int, n_init, p0, p1, alpha)

  # Futility
  if (cp_obs < cp_futility)
    return(c(success = 0L, n = n_int, zone = 0L, n_cap = 0L))

  # Favorable, no extension
  if (cp_obs >= cp_upper) {
    n_final <- n_init
    zone    <- 3L

  # Promising, Mehta–Pocock SSR: extend N until CP rises back to cp_upper
  } else if (cp_obs >= cp_lower) {
    n_final <- ssr_cp_fn(x_int, n_int, p0, p1, alpha,
                         cp_upper, n_init, n_max)
    zone    <- 2L

  # Unfavorable, continue, no increase
  } else {
    n_final <- n_init
    zone    <- 1L
  }

  x_rem    <- rbinom(1L, n_final - n_int, p_true)
  x_total  <- x_int + x_rem

  # Final analysis: one-sided one-proportion z-test (normal approximation),
  # which is the same critical count cp_fn() uses when computing CP. The
  # exact binomial test is uniformly conservative at fixed N and would
  # mechanically suppress T1E below alpha, masking the SSR-induced
  # inflation; the z-test sits close to nominal alpha and lets the
  # SSR zone-boundary effects show through.
  phat <- x_total / n_final
  z    <- (phat - p0) / sqrt(p0 * (1 - p0) / n_final)
  pval <- 1 - pnorm(z)

  c(success = as.integer(pval < alpha),
    n       = n_final,
    zone    = zone,
    n_cap   = as.integer(n_final == n_max))
}

run_cp_scenario <- function(p_true, cp_lower, n_sim = N_SIM, ...) {
  # Wrap the call so `...` is captured from THIS frame (run_cp_scenario's),
  # not from replicate()'s internal `function(...) expr` substitution, which
  # would otherwise inject the sapply iteration value (0L) as the third
  # positional arg, silently overriding cp_upper.
  one_trial <- function() sim_trial_cp(p_true, cp_lower, ...)
  mat <- t(replicate(n_sim, one_trial()))
  tibble(
    p_true   = p_true,
    cp_lower = cp_lower,
    power    = mean(mat[, "success"]),
    exp_n    = mean(mat[, "n"]),
    pr_fut   = mean(mat[, "zone"] == 0),
    pr_unf   = mean(mat[, "zone"] == 1),
    pr_prom  = mean(mat[, "zone"] == 2),
    pr_fav   = mean(mat[, "zone"] == 3),
    pr_cap   = mean(mat[, "n_cap"]),
    n_p10    = quantile(mat[, "n"], 0.10),
    n_p50    = quantile(mat[, "n"], 0.50),
    n_p90    = quantile(mat[, "n"], 0.90),
  )
}
```


## Bayesian PP SSR

The Bayesian design uses *separate* thresholds for the interim efficacy
stop ($\gamma_\text{int}$) and the final analysis success criterion
($\gamma_\text{final}$). The SSR trigger is the predictive probability
of eventual success at the planned $N$, computed *under*
$\gamma_\text{final}$:

$$
\text{PP}(x_\text{int}, n_\text{int}, N) =
\sum_{x_r=0}^{N - n_\text{int}}
\text{BetaBin}(x_r;\; n_r,\; a_\text{post},\; b_\text{post})
\cdot \mathbf{1}\!\left[P(p > p_0 \mid x_\text{int}+x_r,\, N) \geq \gamma_\text{final}\right] .
$$

At the interim, the trial stops early for efficacy when the posterior
$P(p > p_0 \mid x_\text{int}, n_\text{int}) \geq \gamma_\text{int}$;
otherwise the PP-based zone is checked. We use the standard PP zones
$\text{PP}_\text{fut} = 0.05$ for futility and $\text{PP}_U = 0.50$
for the favorable boundary, with the SSR rule extending $N$ until PP
rises to $\text{PP}_U$ (capped at $N_\text{max}$). $\gamma_\text{int}$
is fixed at $0.99$ throughout (a high bar by design, the alternative
$\gamma_\text{int} = 0.95$ would itself produce
$\Pr(\text{interim efficacy stop} \mid p = p_0) \approx 5.5\%$ at
$n_\text{int} = 65$, exceeding $\alpha$ before the SSR branch even runs).
We sweep $\gamma_\text{final}$ over $\{0.950, 0.955, \ldots, 0.990\}$.
The final analysis declares success when the posterior at $n_\text{final}$
exceeds $\gamma_\text{final}$.

```{r bayes-trial-sim}
#| eval: false
# Loaded from R/analysis.R; kept here as inline documentation.
sim_trial_bayes <- function(p_true, gamma_fin, gamma_int = 0.99,
                             pp_fut = 0.05, pp_upper = 0.50,
                             n_init = 84L, n_int = 65L, n_max = 200L,
                             p0 = 0.23, ap = 0.5, bp = 0.5) {
  x_int <- rbinom(1L, n_int, p_true)

  # ── Interim efficacy stop: posterior P(p > p0 | x_int) >= gamma_int ──
  # gamma_int is independent of gamma_final (the final-analysis bar).
  post_int <- post_prob(x_int, n_int, p0, ap, bp)
  if (post_int >= gamma_int) {
    return(c(success = 1L, n = n_int, zone = 3L,
             pr_eff = 1L, stop_fut = 0L, n_cap = 0L))
  }

  # Predictive probability of eventual success, computed under gamma_final
  pp <- pred_prob_fn(x_int, n_int, n_init, p0, ap, bp, gamma_fin)

  # Futility stop
  if (pp <= pp_fut)
    return(c(success = 0L, n = n_int, zone = 0L,
             pr_eff = 0L, stop_fut = 1L, n_cap = 0L))

  # Zone: favorable vs. promising (SSR)
  if (pp >= pp_upper) {
    n_final <- n_init; zone <- 2L
  } else {
    n_final <- ssr_bayes_fn(x_int, n_int, p0, ap, bp, gamma_fin,
                            pp_upper, n_init, n_max)
    zone    <- 1L
  }

  x_rem    <- rbinom(1L, n_final - n_int, p_true)
  x_total  <- x_int + x_rem
  post_fin <- post_prob(x_total, n_final, p0, ap, bp)

  c(success  = as.integer(post_fin >= gamma_fin),
    n        = n_final,
    zone     = zone,
    pr_eff   = 0L,
    stop_fut = 0L,
    n_cap    = as.integer(n_final == n_max))
}

run_bayes_scenario <- function(p_true, gamma_fin, n_sim = N_SIM, ...) {
  # Same `...`-injection guard as run_cp_scenario: replicate()'s internal
  # `function(...) expr` would otherwise pass the sapply iteration value
  # as the third positional arg, silently setting gamma_int = 0.
  one_trial <- function() sim_trial_bayes(p_true, gamma_fin, ...)
  mat <- t(replicate(n_sim, one_trial()))
  tibble(
    p_true    = p_true,
    gamma_fin = gamma_fin,
    power     = mean(mat[, "success"]),
    exp_n     = mean(mat[, "n"]),
    pr_eff    = mean(mat[, "pr_eff"]),
    pr_fut    = mean(mat[, "stop_fut"]),
    pr_prom   = mean(mat[, "zone"] == 1),
    pr_fav    = mean(mat[, "zone"] == 2),
    pr_cap    = mean(mat[, "n_cap"]),
    n_p10     = quantile(mat[, "n"], 0.10),
    n_p50     = quantile(mat[, "n"], 0.50),
    n_p90     = quantile(mat[, "n"], 0.90),
  )
}
```



## Exact Operating Characteristics

Because each design has a single interim look and a binomial endpoint,
the joint distribution of $(x_\text{int}, x_\text{rem})$ under any
true response rate $p$ is finite and discrete. We can therefore compute
all operating characteristics, Type I error, power, expected sample
size, zone probabilities, sample-size quantiles, by **exact enumeration**
rather than Monte Carlo. The functions below sum over every reachable
$(x_\text{int}, x_\text{rem})$ pair and return deterministic OC values
with no Monte Carlo noise. We use them for every headline result in
Section 4; the simulator-based `run_*_scenario` functions remain in
the document for users extending to settings where exact enumeration
is awkward (e.g., multiple interim looks).

```{r oc-exact}
#| eval: false
# Loaded from R/analysis.R; kept here as inline documentation.

# ── Fixed-N z-test Type I error (no interim, no SSR) ─────────────────────
fixed_n_t1e <- function(n, p0 = 0.23, alpha = 0.05) {
  xc <- x_crit_z_fn(n, p0, alpha)
  if (xc > n) return(0)
  1 - pbinom(xc - 1L, n, p0)
}

# ── Helper: exact quantile of a discrete distribution (n_final, prob) ────
discrete_quantile <- function(values, probs, q) {
  ord <- order(values)
  v   <- values[ord]
  cp  <- cumsum(probs[ord])
  v[which(cp >= q)[1]]
}

# ── Exact OC for the CP promising-zone design ────────────────────────────
# Sums over x_int = 0..n_int. For each x_int we deterministically resolve
# zone -> n_final, then compute P(reject final | x_int, n_final, p_true)
# from the binomial CDF of x_rem. No Monte Carlo.
cp_oc_exact <- function(p_true, cp_lower,
                        cp_upper = 0.80, cp_futility = 0.10,
                        n_init = 84L, n_int = 65L, n_max = 200L,
                        p0 = 0.23, p1 = 0.35, alpha = 0.05,
                        ap = 0.5, bp = 0.5) {
  x_grid    <- 0L:n_int
  pmf_x_int <- dbinom(x_grid, n_int, p_true)
  cp_obs    <- vapply(x_grid, function(x) cp_fn(x, n_int, n_init, p0, p1, alpha),
                      numeric(1))

  zone <- integer(length(x_grid))
  zone[cp_obs <  cp_futility]                       <- 0L  # futility
  zone[cp_obs >= cp_upper]                          <- 3L  # favorable
  zone[cp_obs >= cp_lower & cp_obs < cp_upper]      <- 2L  # promising
  zone[cp_obs >= cp_futility & cp_obs < cp_lower]   <- 1L  # unfavorable

  n_final <- ifelse(zone == 0L, n_int,
              ifelse(zone == 2L,
                vapply(x_grid, function(x) ssr_cp_fn(x, n_int, p0, p1, alpha,
                                                      cp_upper, n_init, n_max),
                       integer(1)),
                n_init))

  p_succ <- vapply(seq_along(x_grid), function(i) {
    if (zone[i] == 0L) return(0)
    xc     <- x_crit_z_fn(n_final[i], p0, alpha)
    x_need <- max(0L, xc - x_grid[i])
    n_rem  <- n_final[i] - n_int
    if (x_need > n_rem) return(0)
    1 - pbinom(x_need - 1L, n_rem, p_true)
  }, numeric(1))

  tibble(
    p_true   = p_true,
    cp_lower = cp_lower,
    power    = sum(pmf_x_int * p_succ),
    exp_n    = sum(pmf_x_int * n_final),
    pr_fut   = sum(pmf_x_int * (zone == 0L)),
    pr_unf   = sum(pmf_x_int * (zone == 1L)),
    pr_prom  = sum(pmf_x_int * (zone == 2L)),
    pr_fav   = sum(pmf_x_int * (zone == 3L)),
    pr_cap   = sum(pmf_x_int * (n_final == n_max)),
    n_p10    = discrete_quantile(n_final, pmf_x_int, 0.10),
    n_p50    = discrete_quantile(n_final, pmf_x_int, 0.50),
    n_p90    = discrete_quantile(n_final, pmf_x_int, 0.90)
  )
}

# ── Exact OC for the Bayesian PP design ──────────────────────────────────
bayes_oc_exact <- function(p_true, gamma_fin, gamma_int = 0.99,
                            pp_fut = 0.05, pp_upper = 0.50,
                            n_init = 84L, n_int = 65L, n_max = 200L,
                            p0 = 0.23, ap = 0.5, bp = 0.5) {
  x_grid    <- 0L:n_int
  pmf_x_int <- dbinom(x_grid, n_int, p_true)

  post_int <- vapply(x_grid, function(x) post_prob(x, n_int, p0, ap, bp),
                     numeric(1))
  pp <- vapply(x_grid, function(x)
                pred_prob_fn(x, n_int, n_init, p0, ap, bp, gamma_fin),
                numeric(1))

  # Zones: 3=interim efficacy, 0=futility, 2=favorable (no SSR), 1=promising (SSR)
  zone <- ifelse(post_int >= gamma_int, 3L,
           ifelse(pp <= pp_fut,         0L,
            ifelse(pp >= pp_upper,      2L, 1L)))

  n_final <- ifelse(zone %in% c(0L, 3L), n_int,
              ifelse(zone == 2L, n_init,
                vapply(x_grid, function(x) ssr_bayes_fn(x, n_int, p0, ap, bp,
                                                         gamma_fin, pp_upper,
                                                         n_init, n_max),
                       integer(1))))

  p_succ <- vapply(seq_along(x_grid), function(i) {
    if (zone[i] == 0L) return(0)         # futility, no success
    if (zone[i] == 3L) return(1)         # interim efficacy, success at n_int
    xc     <- x_crit_fn(n_final[i], p0, ap, bp, gamma_fin)
    x_need <- max(0L, xc - x_grid[i])
    n_rem  <- n_final[i] - n_int
    if (x_need > n_rem) return(0)
    1 - pbinom(x_need - 1L, n_rem, p_true)
  }, numeric(1))

  tibble(
    p_true    = p_true,
    gamma_fin = gamma_fin,
    power     = sum(pmf_x_int * p_succ),
    exp_n     = sum(pmf_x_int * n_final),
    pr_eff    = sum(pmf_x_int * (zone == 3L)),
    pr_fut    = sum(pmf_x_int * (zone == 0L)),
    pr_prom   = sum(pmf_x_int * (zone == 1L)),
    pr_fav    = sum(pmf_x_int * (zone == 2L)),
    pr_cap    = sum(pmf_x_int * (n_final == n_max)),
    n_p10     = discrete_quantile(n_final, pmf_x_int, 0.10),
    n_p50     = discrete_quantile(n_final, pmf_x_int, 0.50),
    n_p90     = discrete_quantile(n_final, pmf_x_int, 0.90)
  )
}

# ── CP-driven SSR with Bayesian posterior final analysis ────────────────
# Same interim/zone/SSR logic as cp_oc_exact, but the final analysis
# declares success when post_prob(x_total, n_final) >= gamma_fin
# (instead of the z-test). This isolates which component of the original
# Bayesian-vs-CP advantage comes from the final analysis rule.
cp_oc_post_final <- function(p_true, cp_lower, gamma_fin = 0.955,
                              cp_upper = 0.80, cp_futility = 0.10,
                              n_init = 84L, n_int = 65L, n_max = 200L,
                              p0 = 0.23, p1 = 0.35, alpha = 0.05,
                              ap = 0.5, bp = 0.5) {
  x_grid    <- 0L:n_int
  pmf_x_int <- dbinom(x_grid, n_int, p_true)
  cp_obs    <- vapply(x_grid, function(x) cp_fn(x, n_int, n_init, p0, p1, alpha),
                      numeric(1))

  zone <- integer(length(x_grid))
  zone[cp_obs <  cp_futility]                     <- 0L
  zone[cp_obs >= cp_upper]                        <- 3L
  zone[cp_obs >= cp_lower & cp_obs < cp_upper]    <- 2L
  zone[cp_obs >= cp_futility & cp_obs < cp_lower] <- 1L

  n_final <- ifelse(zone == 0L, n_int,
              ifelse(zone == 2L,
                vapply(x_grid, function(x) ssr_cp_fn(x, n_int, p0, p1, alpha,
                                                      cp_upper, n_init, n_max),
                       integer(1)),
                n_init))

  # Final rule: posterior P(p > p0 | x_total, n_final) >= gamma_fin.
  # Equivalently, x_total >= x_crit_fn(n_final, p0, ap, bp, gamma_fin).
  p_succ <- vapply(seq_along(x_grid), function(i) {
    if (zone[i] == 0L) return(0)
    xc     <- x_crit_fn(n_final[i], p0, ap, bp, gamma_fin)
    if (xc > n_final[i]) return(0)
    x_need <- max(0L, xc - x_grid[i])
    n_rem  <- n_final[i] - n_int
    if (x_need > n_rem) return(0)
    1 - pbinom(x_need - 1L, n_rem, p_true)
  }, numeric(1))

  tibble(
    p_true    = p_true,
    cp_lower  = cp_lower,
    gamma_fin = gamma_fin,
    power     = sum(pmf_x_int * p_succ),
    exp_n     = sum(pmf_x_int * n_final),
    pr_fut    = sum(pmf_x_int * (zone == 0L)),
    pr_unf    = sum(pmf_x_int * (zone == 1L)),
    pr_prom   = sum(pmf_x_int * (zone == 2L)),
    pr_fav    = sum(pmf_x_int * (zone == 3L)),
    pr_cap    = sum(pmf_x_int * (n_final == n_max))
  )
}
```


## Computation Approach

Because each design has a single interim look and a binomial endpoint,
every operating characteristic, Type I error, power, expected sample
size, zone probabilities, sample-size quantiles, is **exactly
computable** by enumerating the joint distribution of
$(x_\text{int}, x_\text{rem})$ under any true response rate $p$. The
helper functions `cp_oc_exact` and `bayes_oc_exact` (Section 3.6) do
this; every numerical result reported below is exact and deterministic
(no Monte Carlo). The simulator-based functions `sim_trial_*` and
`run_*_scenario` remain in the document for users who need to extend
to settings where exact enumeration is awkward (e.g., multiple interim
looks); a fixed seed (`set.seed(20260414)`) makes them reproducible.

Six true ORR scenarios are evaluated: $p_0$, three intermediate values
($p_0 + 0.25\Delta$, $p_0 + 0.50\Delta$, $p_0 + 0.75\Delta$ with
$\Delta = p_1 - p_0$), $p_1$, and an over-performing arm
$p_1 + 0.25\Delta$. For each design we report power (or Type I error
when $p = p_0$), expected sample size $E[N]$, zone probabilities,
budget-cap probability, and the $p10$/$p50$/$p90$ quantiles of
$n_\text{final}$. Generalizability is assessed by repeating the
CP $\text{CP}_L$ sweep across five distinct $(p_0, p_1, n_\text{int},
N_\text{init}, N_\text{max})$ settings. All code is in R 4.x with
Quarto; the manuscript is fully reproducible from this `.qmd`.


# Results


## CP Promising Zone: Type I Error Above $\alpha$ Across the Grid

@fig-cp-t1e is the central empirical result for the base setting. Exact
Type I error at $p = p_0$ sits **uniformly above** the nominal
$\alpha = 0.05$ as $\text{CP}_L$ is raised from $0.30$ to $0.80$,
varying only mildly across the grid. The variation is small but
deterministic: adjacent $\text{CP}_L$ values often map the same
interim integer counts into the same zones, so the exact T1E curve
changes only in steps when an integer crosses a zone boundary.
**No $\text{CP}_L$ on the grid achieves $\text{T1E} \leq \alpha$ in
this setting.** @tbl-cp-oc below reports
power at $p = p_1$ alongside T1E at each $\text{CP}_L$, power
decreases gently as the promising zone shrinks. Section 4.2 traces
the inflation to the discrete arithmetic of the $z$-test critical
count combined with SSR-driven changes in $N$; Section 4.5 then
shows that the *direction* of the bias changes from one design
configuration to another, so the base-setting inflation is not
universal.

```{r cp-sweep}
#| cache: true
#| message: false

# Headline OC: exact enumeration (no Monte Carlo noise).
cp_sweep <- expand_grid(
  cp_lower = cp_lower_grid,
  p_true   = scenarios
) |>
  pmap_dfr(function(cp_lower, p_true) {
    cp_oc_exact(p_true = p_true, cp_lower = cp_lower)
  })
```

```{r cp-t1e-plot}
#| label: fig-cp-t1e
#| fig-cap: "Exact Type I error as a function of the CP promising lower bound, zoomed on the 4–8% range. The T1E curve sits uniformly above the nominal $\\alpha$ = 0.05 across every $\\text{CP}_L$ on the grid, the key finding. (Power at $p_1$ is shown in the head-to-head comparison, §4.4.)"

t1e_only <- cp_sweep |>
  filter(p_true == p0)

ggplot(t1e_only, aes(x = cp_lower, y = power)) +
  geom_hline(yintercept = alpha, color = "red", linewidth = 0.5,
             linetype = "dotted") +
  geom_line(color = "#C0392B", linewidth = 1.2) +
  geom_point(color = "#C0392B", size = 3, shape = 21,
             fill = "white", stroke = 1.5) +
  annotate("text", x = 0.325, y = alpha + 0.0015,
           label = glue("alpha = {alpha}"), size = 3, color = "red") +
  scale_x_continuous(breaks = cp_lower_grid,
                     labels = as.character(cp_lower_grid)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1),
                     breaks = seq(0.04, 0.08, 0.005),
                     limits = c(0.04, 0.08)) +
  labs(
    title    = "CP Promising Zone SSR: Type I Error vs. Promising Lower Bound",
    subtitle = TeX(glue(
      "$p_0 = {p0}$, $p_1 = {p1}$, $n_{{int}} = {n_int}$, ",
      "$N_{{init}} = {n_init}$, $N_{{max}} = {n_max}$, ",
      "$CP_{{upper}} = {cp_upper}$, $CP_{{fut}} = {cp_futility}$, exact enumeration"
    )),
    x = TeX("CP Promising Lower ($CP_L$)"),
    y = TeX("Exact Type I Error ($p = p_0$)")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", color = "#1F4E79", size = 11),
    plot.subtitle    = element_text(color = "gray40", size = 9),
    panel.grid.minor = element_blank()
  )
```

@tbl-cp-oc reports the full operating characteristics at six representative
$\text{CP}_L$ values. The Type I error inflation is small in absolute
terms, typically a few tenths of a percentage point above $\alpha$,
but it is robust: every $\text{CP}_L$ on the grid produces a
nominally-uncalibrated design, and lower $\text{CP}_L$ values trade
slightly higher T1E for somewhat higher power at $p_1$.

```{=latex}
\begin{landscape}
```

```{r cp-oc-table}
#| label: tbl-cp-oc
#| tbl-cap: "CP SSR operating characteristics at selected $\\text{CP}_L$ values. Power / T1E column reports exact power at the corresponding True ORR; at $p = p_0$ = 23% it is the exact Type I error (target ≤ 5%). Design parameters and computation method appear in the table subtitle."
selected_pl <- c(0.30, 0.50, 0.65, 0.80)

# Use round() to avoid floating-point mismatches from seq()
cp_sweep |>
  filter(round(cp_lower, 4) %in% round(selected_pl, 4)) |>
  left_join(scen_label_tbl, by = "p_true") |>
  mutate(
    `CP_L`          = round(cp_lower, 2),
    `True ORR`      = percent(p_true, accuracy = 1),
    `Power / T1E`   = percent(power,   accuracy = 0.1),
    `E[N]`          = round(exp_n, 1),
    `Pr(Fut)`       = percent(pr_fut,  accuracy = 1),
    `Pr(Prom)`      = percent(pr_prom, accuracy = 1),
    `Pr(Fav)`       = percent(pr_fav,  accuracy = 1),
    `Pr(Cap)`       = percent(pr_cap,  accuracy = 1),
    `N p10/p50/p90` = paste0(n_p10, " / ", n_p50, " / ", n_p90)
  ) |>
  arrange(`CP_L`, p_true) |>
  select(`CP_L`, `True ORR`, Label, `Power / T1E`, `E[N]`,
         `Pr(Fut)`, `Pr(Prom)`, `Pr(Fav)`, `Pr(Cap)`, `N p10/p50/p90`) |>
  gt(groupname_col = "CP_L") |>
  tab_source_note(md(glue(
    "Design: $\\text{{CP}}_\\text{{upper}} = {cp_upper}$ · ",
    "$\\text{{CP}}_\\text{{fut}} = {cp_futility}$ · ",
    "$n_\\text{{int}} = {n_int}$ · $N_\\text{{max}} = {n_max}$ · ",
    "exact enumeration."
  ))) |>
  tab_style(
    style    = list(cell_fill(color = "#F8D7DA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "23%")
  ) |>
  tab_style(
    style    = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "35%")
  ) |>
  tab_footnote(
    "Power / T1E column: at p = p\u2080 this is the exact Type I error; target \u2264 5.0%.",
    locations = cells_column_labels(columns = `Power / T1E`)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              row_group.font.weight = "bold",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{=latex}
\end{landscape}
```


## Mechanism: Discrete $z$-test Bias and Boundary Crossings

The Type I error pattern has two sources, both arising from the
discrete one-sided $z$-test critical count $x_\text{crit}(N)$ used at
the final analysis. First, at any finite $N$, $x_\text{crit}(N)$ is
the smallest integer with $z \geq z_\alpha$, so the actual fixed-$N$
rejection probability under $H_0$ differs from the nominal $\alpha$
by a *signed* amount that depends on where the rounding falls, in
some $(p_0, N)$ configurations the rounding leaves T1E above $\alpha$,
in others below. In our base setting ($p_0 = 0.23$, $N_\text{init} = 84$)
the bias is upward; Section 4.5 documents settings where it is
downward. Second, SSR moves trials between finite-$N$ tests whose
biases can have either sign, and the per-integer analysis below shows
the *direction* of the SSR contribution at each boundary crossing.
At $n_\text{int} = 65$, $x_\text{int}$ takes $66$ possible values;
each maps to a specific CP at $N_\text{init}$. As $\text{CP}_L$ is
raised from one grid point to the next, only the integers whose CP
falls in that interval cross from the **promising** zone into the
**unfavorable** zone. Each such crossing changes that integer's
contribution from "extend $N$ via SSR" to "stay at $N_\text{init}$".

The signed contribution to the Type I error when an integer leaves the
promising zone is
$\Delta\text{T1E}(x_\text{int})
= \Pr(x_\text{int} \mid p_0)
\cdot \bigl[\text{FP}_\text{init}(x_\text{int})
- \text{FP}_\text{ssr}(x_\text{int})\bigr]$,
where $\text{FP}_\text{init}$ is the conditional false-positive
probability if the trial stays at $N_\text{init}$ and $\text{FP}_\text{ssr}$
is the false-positive probability if SSR drives $N$ to its (per-integer)
target. In principle, because the discrete $z$-test critical count is not
exactly $\alpha$ at any finite $N$, the sign of $\Delta\text{T1E}$
could go either way at each crossing. **In our base setting all
observed crossings have $\Delta\text{T1E} < 0$**, so exact T1E
declines monotonically (and only mildly) as $\text{CP}_L$ rises,
but never by enough to reach $\alpha$, because the underlying
fixed-$N$ discreteness bias dominates the SSR effect. The stepwise
changes are small because adjacent $\text{CP}_L$ values often
classify the same integer counts into the same zone, so only one
or two integers cross at each grid step.

```{r cp-mechanism}
# ── For each x_int, compute CP at n_init and the SSR target N ─────────────
# n_ssr(x_int) is the smallest n in [n_init, n_max] with CP(n) >= cp_upper,
# i.e., the N this integer would drive to under the SSR rule if it fell in
# the promising zone. fp_at_ssr is the resulting conditional false-positive
# probability at that N. fp_at_init is the FP probability if no SSR were
# done (integer falls in unfavorable). The signed difference governs the
# T1E contribution change when that integer crosses promising -> unfavorable.
mechanism_tbl <- tibble(x_int = 0L:n_int) |>
  mutate(
    cp_at_init  = map_dbl(x_int, ~ cp_fn(.x, n_int, n_init, p0, p1, alpha)),
    prob_x_null = dbinom(x_int, n_int, p0),
    # Critical count under z-test, same helper used by cp_fn / sim_trial_cp
    xc_init     = x_crit_z_fn(n_init, p0, alpha),
    x_need_init = pmax(0L, xc_init - x_int),
    fp_at_init  = map2_dbl(x_int, x_need_init,
                           ~ if (.y > n_init - n_int) 0
                             else 1 - pbinom(.y - 1L, n_init - n_int, p0)),
    # SSR target N per x_int (capped at n_max if cp_upper unreachable)
    n_ssr       = map_int(x_int, function(x) {
      for (nc in seq(n_init, n_max)) {
        if (cp_fn(x, n_int, nc, p0, p1, alpha) >= cp_upper) return(nc)
      }
      n_max
    }),
    xc_ssr      = map_int(n_ssr, ~ x_crit_z_fn(.x, p0, alpha)),
    x_need_ssr  = pmax(0L, xc_ssr - x_int),
    fp_at_ssr   = pmap_dbl(list(x_int, x_need_ssr, n_ssr),
                           function(xi, xn, ns) {
                             nr <- ns - n_int
                             if (xn > nr) 0
                             else 1 - pbinom(xn - 1L, nr, p0)
                           }),
    # T1E contribution change as this integer crosses promising -> unfavorable
    # (loses the SSR-conditional FP, gains the no-SSR FP at n_init)
    t1e_change  = prob_x_null * (fp_at_init - fp_at_ssr)
  )

# ── For each CP_L step, identify integers that cross the boundary ─────────
# When CP_L is raised from pl_old to pl_new (pl_new > pl_old),
# integers with pl_old <= CP < pl_new move from promising → unfavorable
boundary_crossings <- map_dfr(
  seq_along(cp_lower_grid[-1]),
  function(i) {
    pl_old <- cp_lower_grid[i]
    pl_new <- cp_lower_grid[i + 1]
    mechanism_tbl |>
      filter(cp_at_init >= pl_old, cp_at_init < pl_new) |>
      mutate(
        pl_from  = pl_old,
        pl_to    = pl_new,
        step     = glue("{pl_old} -> {pl_new}")
      )
  }
)
```

```{r mechanism-plot}
#| label: fig-mechanism
#| fig-cap: "For each integer response count $x_\\text{int}$ that crosses from promising to unfavorable as $\\text{CP}_L$ is raised, we show its individual Type I error contribution change. Positive values = T1E increases; negative = T1E decreases. Steps with no crossing counts are omitted."

# Only plot steps that have at least one crossing
bc_plot <- boundary_crossings |>
  filter(!is.na(step))

if (nrow(bc_plot) == 0) {
  message("No boundary crossings found, check cp_lower_grid and cp_at_init values.")
} else {
  bc_plot |>
    ggplot(aes(x = factor(x_int), y = t1e_change * 1000,
               fill = t1e_change > 0)) +
    geom_col(width = 0.7, alpha = 0.85) +
    geom_hline(yintercept = 0, linewidth = 0.5) +
    facet_wrap(~ step, scales = "free_x", ncol = 3) +
    scale_fill_manual(
      values = c("TRUE" = "#C0392B", "FALSE" = "#2E75B6"),
      labels = c("TRUE" = "T1E increases", "FALSE" = "T1E decreases"),
      name   = NULL
    ) +
    labs(
      title    = "Type I Error Contribution Change at Each Boundary Crossing",
      subtitle = TeX(paste0(
        "Each panel = one $CP_L$ step. ",
        "Bars = $x_{int}$ values crossing from promising $\\to$ ",
        "unfavorable zone. Y-axis: $\\Delta T1E \\times 10^3$."
      )),
      x = TeX("Interim response count ($x_{int}$)"),
      y = TeX("$\\Delta T1E \\times 10^3$")
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.title       = element_text(face = "bold", color = "#1F4E79", size = 11),
      plot.subtitle    = element_text(color = "gray40", size = 9),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom",
      strip.text       = element_text(face = "bold", size = 9)
    )
}
```

```{r mechanism-table}
#| label: tbl-mechanism
#| tbl-cap: "Summary of T1E contribution changes at each $\\text{CP}_L$ step. In our base setting every observed crossing has $\\Delta$T1E < 0, so exact T1E declines monotonically as $\\text{CP}_L$ rises, but the magnitudes are small and the underlying fixed-N discreteness bias keeps T1E above $\\alpha$ at every $\\text{CP}_L$. The 'Exact T1E' column reports the deterministic value from cp_oc_exact at the new $\\text{CP}_L$. Steps with no crossings mean no $x_\\text{int}$ values fall in that CP interval, the discrete distribution has a gap there."

if (nrow(boundary_crossings) > 0) {
  boundary_crossings |>
    group_by(step, pl_from, pl_to) |>
    summarise(
      x_values       = paste(sort(x_int), collapse = ", "),
      net_t1e_change = sum(t1e_change),
      .groups        = "drop"
    ) |>
    left_join(
      cp_sweep |>
        filter(p_true == p0) |>
        select(cp_lower, t1e = power),
      by = c("pl_to" = "cp_lower")
    ) |>
    mutate(
      `CP_L step`        = step,
      `Counts crossing`  = x_values,
      `Net dT1E`         = sprintf("%+.4f", net_t1e_change),
      `Direction`        = if_else(net_t1e_change > 0, "Up", "Down"),
      `Exact T1E`        = percent(t1e, accuracy = 0.01)
    ) |>
    select(`CP_L step`, `Counts crossing`,
           `Net dT1E`, `Direction`, `Exact T1E`) |>
    gt() |>
    cols_label(
      `CP_L step`       = md("$\\text{CP}_L$ step"),
      `Counts crossing` = "Counts crossing",
      `Net dT1E`        = md("Net $\\Delta$T1E"),
      `Direction`       = "Direction",
      `Exact T1E`       = "Exact T1E"
    ) |>
    tab_source_note(md(glue("Design: $\\text{{CP}}_\\text{{upper}} = {cp_upper}$, $\\text{{CP}}_\\text{{fut}} = {cp_futility}$, $n_\\text{{int}} = {n_int}$, $N_\\text{{init}} = {n_init}$."))) |>
    tab_style(
      style    = list(cell_fill(color = "#F8D7DA"), cell_text(weight = "bold")),
      locations = cells_body(rows = `Direction` == "Up")
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels()
    ) |>
    tab_options(table.width = pct(100), table.font.size = "small",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
} else {
  cat("No boundary crossings detected. Inspect mechanism_tbl$cp_at_init distribution.")
}
```


## Bayesian PP SSR: Monotone Calibration

@fig-bayes-t1e shows the Bayesian counterpart of @fig-cp-t1e. Exact Type I
error decreases **monotonically** in $\gamma_\text{final}$; power
likewise decreases monotonically. Both curves are still stepwise
because the final success rule reduces to an integer-grid decision
($x_\text{total} \geq x_\text{crit}(n_\text{final}, \gamma_\text{final})$),
so adjacent $\gamma_\text{final}$ values that share the same integer
critical count produce identical T1E, the rendered table below
shows visible plateaus across consecutive grid points. The
calibration surface is nonetheless **far less irregular** than the
$\text{CP}_L$ surface in §4.1: PP integrates over a continuous
posterior, so as $\gamma_\text{final}$ rises the integer critical
count also rises (or stays the same), monotonically reducing T1E.
The practical consequence is that the design can be calibrated to
any target Type I error by enumeration, simply by
choosing the smallest $\gamma_\text{final}$ that meets the Type I
error budget.

```{r bayes-sweep}
#| cache: true

# Headline OC: exact enumeration (no Monte Carlo noise).
bayes_sweep <- expand_grid(
  gamma_fin = gamma_fin_grid,
  p_true    = scenarios
) |>
  pmap_dfr(function(gamma_fin, p_true) {
    bayes_oc_exact(p_true = p_true, gamma_fin = gamma_fin)
  })
```

```{r bayes-t1e-plot}
#| label: fig-bayes-t1e
#| fig-cap: "Exact Type I error (solid) and power at $p_1$ (dashed) as a function of $\\gamma_\\text{final}$ on the evaluated grid. Smooth, monotone behaviour contrasts sharply with the CP result."

t1e_power_bayes <- bayes_sweep |>
  filter(p_true %in% c(p0, p1)) |>
  mutate(metric = if_else(p_true == p0,
                          "Type I Error (p = p0)", "Power (p = p1)"))

ggplot(t1e_power_bayes, aes(x = gamma_fin, y = power,
                             color = metric, linetype = metric)) +
  geom_hline(yintercept = alpha, color = "red",   linewidth = 0.5, linetype = "dotted") +
  geom_hline(yintercept = power, color = "gray50", linewidth = 0.5, linetype = "dotted") +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3, shape = 21, fill = "white", stroke = 1.5) +
  annotate("text", x = 0.985, y = alpha + 0.03,
           label = glue("alpha = {alpha}"), size = 3, color = "red") +
  annotate("text", x = 0.985, y = power + 0.03,
           label = "Target power = 0.80", size = 3, color = "gray40") +
  scale_color_manual(values = c("Type I Error (p = p0)" = "#C0392B",
                                "Power (p = p1)"        = "#2E75B6")) +
  scale_linetype_manual(values = c("Type I Error (p = p0)" = "solid",
                                   "Power (p = p1)"        = "dashed")) +
  scale_x_continuous(breaks = seq(0.95, 0.99, by = 0.01),
                     labels = as.character(seq(0.95, 0.99, by = 0.01))) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     breaks = seq(0, 1, 0.10),
                     limits = c(0, NA)) +
  labs(
    title    = TeX("Bayesian PP SSR: Type I Error and Power vs. $\\gamma_{final}$"),
    subtitle = TeX(glue(
      "$p_0 = {p0}$, $p_1 = {p1}$, $n_{{int}} = {n_int}$, $N_{{max}} = {n_max}$, ",
      "$\\gamma_{{int}} = {gamma_int}$, $PP_{{upper}} = 0.50$, exact enumeration"
    )),
    x = TeX("$\\gamma_{final}$ (final analysis posterior threshold)"),
    y = "Exact Probability",
    color = NULL, linetype = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", color = "#1F4E79", size = 11),
    plot.subtitle    = element_text(color = "gray40", size = 9),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom"
  )
```

```{=latex}
\begin{landscape}
```

```{r bayes-oc-table}
#| label: tbl-bayes-oc
#| tbl-cap: "Bayesian PP SSR operating characteristics at selected $\\gamma_\\text{final}$ values. Power / T1E column reports exact power at the corresponding True ORR; at $p = p_0$ = 23% it is the exact Type I error (target ≤ 5%). Design parameters and prior specification appear in the table subtitle."
selected_gf <- c(0.95, 0.96, 0.975, 0.99)

bayes_sweep |>
  filter(round(gamma_fin, 4) %in% round(selected_gf, 4)) |>
  left_join(scen_label_tbl, by = "p_true") |>
  mutate(
    `gamma_final`   = round(gamma_fin, 3),
    `True ORR`      = percent(p_true, accuracy = 1),
    `Power / T1E`   = percent(power,   accuracy = 0.1),
    `E[N]`          = round(exp_n, 1),
    `Pr(Eff stop)`  = percent(pr_eff,  accuracy = 1),
    `Pr(Fut)`       = percent(pr_fut,  accuracy = 1),
    `Pr(Prom)`      = percent(pr_prom, accuracy = 1),
    `Pr(Fav)`       = percent(pr_fav,  accuracy = 1),
    `Pr(Cap)`       = percent(pr_cap,  accuracy = 1),
    `N p10/p50/p90` = paste0(n_p10, " / ", n_p50, " / ", n_p90)
  ) |>
  arrange(gamma_final, p_true) |>
  select(`gamma_final`, `True ORR`, Label, `Power / T1E`, `E[N]`,
         `Pr(Eff stop)`, `Pr(Fut)`, `Pr(Prom)`, `Pr(Fav)`,
         `Pr(Cap)`, `N p10/p50/p90`) |>
  gt(groupname_col = "gamma_final") |>
  tab_source_note(md(glue(
    "Design: $\\gamma_\\text{{int}} = {gamma_int}$ · ",
    "$\\text{{PP}}_\\text{{upper}} = 0.50$ · ",
    "$n_\\text{{int}} = {n_int}$ · $N_\\text{{max}} = {n_max}$ · ",
    "Prior $\\text{{Beta}}({ap},{bp})$ · exact enumeration."
  ))) |>
  tab_style(
    style    = list(cell_fill(color = "#F8D7DA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "23%")
  ) |>
  tab_style(
    style    = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "35%")
  ) |>
  tab_footnote(
    "Power / T1E at p = p\u2080 is the exact Type I error; target \u2264 5.0%.",
    locations = cells_column_labels(columns = `Power / T1E`)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              row_group.font.weight = "bold",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{=latex}
\end{landscape}
```


## Head-to-Head Comparison

For a head-to-head comparison we calibrate each design directly from
the exact-enumeration sweeps. The Bayesian design is calibrated to
the smallest $\gamma_\text{final}$ on the grid with exact
T1E $\leq \alpha$; because Bayesian T1E is monotone, this is the
highest-power T1E-controlling design. The CP design is calibrated to
the $\text{CP}_L$ value with the highest exact power at $p_1$ among
those with T1E $\leq \alpha$ (and flagged as uncalibratable if no
$\text{CP}_L$ on the grid meets the constraint). The summary below
is computed from the exact outputs above, not from any external
benchmark.

```{r calibration-summary}
#| label: tbl-calibration
#| tbl-cap: "Calibration summary comparing the CP and Bayesian PP designs. The CP design is uncalibratable in the base setting because every $\\text{CP}_L$ on the grid produces exact T1E above $\\alpha$ = 5% (see Table 1); the Bayesian PP design calibrates at the smallest $\\gamma_\\text{final}$ with exact T1E ≤ $\\alpha$. Calibration values are deterministic (exact enumeration, no Monte Carlo)."
# Bayesian: smallest gamma_final on the grid with T1E <= alpha.
# T1E is monotone-decreasing in gamma_fin, so this is the highest-power
# Bayesian design that controls Type I error.
bayes_t1e_curve <- bayes_sweep |> filter(p_true == p0) |> arrange(gamma_fin)

bayes_cal_row <- bayes_t1e_curve |>
  filter(power <= alpha) |>
  slice_min(order_by = gamma_fin, n = 1)

gamma_fin_cal <- if (nrow(bayes_cal_row) > 0) {
  bayes_cal_row$gamma_fin
} else {
  max(gamma_fin_grid)
}

bayes_t1e <- bayes_t1e_curve |>
  filter(round(gamma_fin, 4) == round(gamma_fin_cal, 4)) |>
  pull(power)

bayes_power <- bayes_sweep |>
  filter(p_true == p1, round(gamma_fin, 4) == round(gamma_fin_cal, 4)) |>
  pull(power)

# CP: among CP_L values with T1E <= alpha, pick the one with highest power
# at p1. (CP T1E drifts only mildly with CP_L and may not be strictly
# monotone in every parameter setting, so we rank by power-at-p1 rather
# than by CP_L itself.)
cp_t1e_curve <- cp_sweep |> filter(p_true == p0) |> select(cp_lower, t1e = power)
cp_pwr_curve <- cp_sweep |> filter(p_true == p1) |> select(cp_lower, pwr = power)
cp_curves    <- cp_t1e_curve |> inner_join(cp_pwr_curve, by = "cp_lower")

cp_under_alpha <- cp_curves |> filter(t1e <= alpha) |> arrange(desc(pwr))

cp_best_pl  <- if (nrow(cp_under_alpha) > 0) cp_under_alpha$cp_lower[1] else NA_real_
cp_t1e      <- if (!is.na(cp_best_pl)) cp_under_alpha$t1e[1]            else NA_real_
cp_power_p1 <- if (!is.na(cp_best_pl)) cp_under_alpha$pwr[1]            else NA_real_

cp_label     <- if (!is.na(cp_best_pl))  glue("$\\text{{CP}}_L = {cp_best_pl}$") else "None"
cp_t1e_label <- if (!is.na(cp_t1e))      percent(cp_t1e,      accuracy = 0.1) else "\u2014"
cp_pwr_label <- if (!is.na(cp_power_p1)) percent(cp_power_p1, accuracy = 0.1) else "\u2014"
cp_cal_label <- if (!is.na(cp_best_pl)) "Yes" else "No"

tribble(
  ~Method, ~`Calibration parameter`, ~`T1E at p0`, ~`Power at p1`, ~`Calibratable`,
  "CP Promising Zone",
    cp_label,
    cp_t1e_label,
    cp_pwr_label,
    cp_cal_label,
  "Bayesian PP",
    glue("$\\gamma_\\text{{final}} = {sprintf('%.3f', gamma_fin_cal)}$"),
    percent(bayes_t1e,   accuracy = 0.1),
    percent(bayes_power, accuracy = 0.1),
    "Yes"
) |>
  gt() |>
  fmt_markdown(columns = `Calibration parameter`) |>
  cols_label(
    Method                  = "Method",
    `Calibration parameter` = "Calibration parameter",
    `T1E at p0`             = md("T1E at $p_0$"),
    `Power at p1`           = md("Power at $p_1$"),
    `Calibratable`          = "Calibratable"
  ) |>
  tab_source_note(md(
    "Calibration values derived from the exact-enumeration sweeps above (deterministic, no Monte Carlo)."
  )) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{r head-to-head}
#| label: tbl-head-to-head
#| tbl-cap: "Head-to-head exact operating characteristics for the calibrated Bayesian PP design and the closest-to-$\\alpha$ CP design across all six true ORR scenarios. Bayesian PP controls T1E and dominates on power; the CP comparator is shown at its closest-to-$\\alpha$ $\\text{CP}_L$ because no $\\text{CP}_L$ on the grid actually achieves T1E ≤ $\\alpha$."
# Run Bayesian at calibrated gamma_fin across all scenarios
bayes_cal <- map_dfr(scenarios, function(pt)
              bayes_oc_exact(p_true = pt, gamma_fin = gamma_fin_cal)) |>
  left_join(scen_label_tbl, by = "p_true") |>
  mutate(method = glue("Bayesian PP (gamma_final = {gamma_fin_cal})"))

# Best CP (or closest)
cp_pl_use <- if (!is.na(cp_best_pl)) {
  cp_best_pl
} else {
  cp_lower_grid[which.min(
    abs(cp_sweep |> filter(p_true == p0) |> pull(power) - alpha)
  )]
}

cp_cal <- map_dfr(scenarios, function(pt)
              cp_oc_exact(p_true = pt, cp_lower = cp_pl_use)) |>
  left_join(scen_label_tbl, by = "p_true") |>
  mutate(method = glue("CP (CP_L = {cp_pl_use})"))

comparison <- bind_rows(bayes_cal, cp_cal)

comparison |>
  transmute(
    Method          = method,
    `True ORR`      = percent(p_true, accuracy = 1),
    Label           = Label,
    `Power / T1E`   = percent(power,  accuracy = 0.1),
    `E[N]`          = round(exp_n, 1),
    `Pr(Fut)`       = percent(pr_fut, accuracy = 1),
    `Pr(Cap)`       = percent(pr_cap, accuracy = 1),
    `N p50 / p90`   = paste0(n_p50, " / ", n_p90)
  ) |>
  gt(groupname_col = "Method") |>
  tab_source_note(md(glue(
    "Design: $n_\\text{{int}} = {n_int}$ · $N_\\text{{init}} = {n_init}$ · ",
    "$N_\\text{{max}} = {n_max}$ · exact enumeration."
  ))) |>
  tab_style(
    style    = list(cell_fill(color = "#F8D7DA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "23%")
  ) |>
  tab_style(
    style    = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `True ORR` == "35%")
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small", row_group.font.weight = "bold",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{r combined-power-plot}
#| label: fig-power-curves
#| fig-cap: "Power curves for the two calibrated designs. Vertical dotted line marks $p_0$; horizontal dashed line marks 80% target."

comparison |>
  filter(p_true > p0) |>
  ggplot(aes(x = p_true, y = power, color = method, linetype = method)) +
  geom_hline(yintercept = 0.80, linetype = "dashed", color = "gray50", linewidth = 0.5) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5, fill = "white", shape = 21, stroke = 1.5) +
  scale_color_manual(values = c("#2E75B6", "#E67E22")) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_x_continuous(labels = percent_format(accuracy = 1),
                     breaks = scenarios[scenarios > p0]) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
  labs(
    title    = "Power Curves: Bayesian PP vs. CP (Calibrated Designs)",
    subtitle = TeX(glue("$n_{{int}} = {n_int}$ · $N_{{max}} = {n_max}$ · exact enumeration")),
    x = "True ORR", y = "Power",
    color = "Design", linetype = "Design"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", color = "#1F4E79", size = 11),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom"
  )
```


## Generalizability

To examine how CP T1E behaviour generalises beyond the base setting,
we repeat the CP $\text{CP}_L$ sweep across five
$(p_0, p_1, n_\text{int}, N_\text{init}, N_\text{max})$ combinations
spanning $p_0 \in \{0.10, 0.20, 0.23, 0.30, 0.40\}$. @fig-generalizability shows
that **CP T1E behaviour is highly design-specific**. In the base
setting ($p_0 = 0.23$) and the $p_0 = 0.20$ setting, exact T1E sits
above $\alpha = 0.05$ at every $\text{CP}_L$ on the grid. In the
$p_0 = 0.10$, $p_0 = 0.30$, and $p_0 = 0.40$ settings, some
$\text{CP}_L$ values produce T1E at or below $\alpha$ while others
produce inflation, with no monotone or otherwise predictable
relationship. The qualitative driver is the same in every setting,
the discrete $z$-test critical count introduces a rounding bias whose
sign and magnitude depend on $(p_0, N)$, and SSR shifts trials between
finite-$N$ tests whose biases vary in sign, but the practical
consequence is that $\text{CP}_L$ cannot be calibrated by a fixed
rule across designs. Each design configuration requires its own
enumeration (or simulation) sweep, and the resulting T1E surface is
too irregular to support a one-size-fits-all calibration.

```{r generalizability}
#| cache: true

gen_grid <- tribble(
  ~p0,  ~p1,  ~n_int, ~n_init, ~n_max, ~label,
  0.10, 0.25,  40L,    55L,    150L,   "p0=0.10, p1=0.25",
  0.20, 0.35,  50L,    72L,    180L,   "p0=0.20, p1=0.35",
  0.23, 0.35,  65L,    84L,    200L,   "p0=0.23, p1=0.35 (base)",
  0.30, 0.45,  55L,    75L,    200L,   "p0=0.30, p1=0.45",
  0.40, 0.55,  60L,    85L,    200L,   "p0=0.40, p1=0.55"
)

# Generalizability sweep: exact enumeration across 5 design configurations.
gen_results <- pmap_dfr(gen_grid, function(p0, p1, n_int, n_init, n_max, label) {
  map_dfr(cp_lower_grid, function(pl) {
    cp_oc_exact(
      p_true   = p0,  # T1E only
      cp_lower = pl,
      n_init   = n_init, n_int = n_int, n_max = n_max,
      p0 = p0, p1 = p1
    ) |>
      mutate(scenario_label = label, p0 = p0, p1 = p1)
  })
})
```

```{r fixed-n-ref}
#| label: tbl-fixed-n
#| tbl-cap: "Fixed-N reference: exact Type I error of the one-sided z-test at the planned $N_\\text{init}$ for each generalizability setting, before any interim or SSR. The signed bias relative to $\\alpha$ = 5% is the rounding bias inherited by the CP design from the discrete final test alone, separate from any SSR contribution. The base setting ($p_0$ = 0.23, N = 84) and the $p_0$ = 0.20 setting are both biased upward; the others are biased downward."

gen_grid |>
  mutate(
    `T1E_fixedN` = pmap_dbl(
      list(p0 = p0, n_init = n_init),
      function(p0, n_init) fixed_n_t1e(n_init, p0, alpha)
    ),
    `Bias vs α` = `T1E_fixedN` - alpha,
    `xc(N_init)` = pmap_int(
      list(n_init = n_init, p0 = p0),
      function(n_init, p0) x_crit_z_fn(n_init, p0, alpha)
    )
  ) |>
  transmute(
    Setting        = label,
    `(p0, p1)`     = sprintf("(%.2f, %.2f)", p0, p1),
    `N_init`       = n_init,
    `xc(N_init)`   = `xc(N_init)`,
    `Fixed-N T1E`  = percent(`T1E_fixedN`, accuracy = 0.01),
    `Bias vs α`    = sprintf("%+.2f%%", `Bias vs α` * 100)
  ) |>
  gt() |>
  cols_label(
    Setting       = "Setting",
    `(p0, p1)`    = md("$(p_0, p_1)$"),
    `N_init`      = md("$N_\\text{init}$"),
    `xc(N_init)`  = md("$x_c(N_\\text{init})$"),
    `Fixed-N T1E` = md("Fixed-$N$ T1E"),
    `Bias vs α`   = md("Bias vs $\\alpha$")
  ) |>
  tab_source_note(md(
    "Rejection probability of the one-sided $z$-test at $N_\\text{init}$ under $H_0: p = p_0$, with no interim or SSR. $\\alpha = 5\\%$ target."
  )) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{r generalizability-plot}
#| label: fig-generalizability
#| fig-cap: "Exact Type I error as a function of $\\text{CP}_L$ across five design parameter settings. CP T1E behaviour is highly design-specific: in some settings T1E sits above $\\alpha$ at every $\\text{CP}_L$; in others, some $\\text{CP}_L$ values control T1E and others do not, with no monotone or otherwise predictable rule. The pattern supports calibration only by per-design enumeration, not by a fixed-$\\text{CP}_L$ rule."

gen_results |>
  ggplot(aes(x = cp_lower, y = power, color = scenario_label)) +
  geom_hline(yintercept = 0.05, linetype = "dotted", color = "red", linewidth = 0.5) +
  geom_line(linewidth = 1.0) +
  geom_point(size = 2.5, shape = 21, fill = "white", stroke = 1.2) +
  annotate("text", x = 0.325, y = 0.046, label = "alpha = 5%",
           size = 3, color = "red") +
  scale_color_brewer(palette = "Dark2", name = "Design setting") +
  scale_x_continuous(breaks = cp_lower_grid) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1),
                     breaks = seq(0.02, 0.10, 0.02),
                     limits = c(0.02, 0.10)) +
  labs(
    title    = "Exact CP T1E Across Design Parameter Settings",
    subtitle = TeX(glue("$CP_{{upper}} = {cp_upper}$, $CP_{{fut}} = {cp_futility}$ · exact enumeration")),
    x = TeX("CP Promising Lower ($CP_L$)"),
    y = "Exact Type I Error"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", color = "#1F4E79", size = 11),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom"
  )
```


## Symmetric Comparison: Does the Final-Analysis Rule Drive the CP Inflation?

The original CP design pairs CP-zone SSR with a one-sided $z$-test
final; the Bayesian design pairs PP-zone SSR with a posterior final.
The asymmetry leaves open whether the Bayesian T1E advantage comes
from the SSR logic, the final-analysis rule, or both. We run a
**partial** symmetric comparison: a single new variant in which the
**CP-driven** SSR is paired with the **Bayesian posterior** final
analysis (same zone classification, same SSR target
$\text{CP}_U = 0.80$, same final threshold $\gamma_\text{final}$
on the posterior $P(p > p_0 \mid x_\text{total}, n_\text{final})$).
The only difference between this variant and the original CP design
is the final analysis rule. If this variant controls T1E, the
final-analysis rule is *sufficient* to explain the original CP
inflation, though without also implementing PP-driven SSR with a
$z$-test final we cannot rule out that the SSR logic *also*
contributes. We leave that complementary variant as future work
(see Discussion).

```{r cp-post-final-sweep}
# CP-driven SSR + posterior final, swept over gamma_fin at the
# best-power CP_L from Section 4.4 (cp_pl_use).
cp_post_sweep <- expand_grid(
  cp_lower  = cp_pl_use,
  gamma_fin = gamma_fin_grid,
  p_true    = c(p0, p1)
) |>
  pmap_dfr(function(cp_lower, gamma_fin, p_true) {
    cp_oc_post_final(
      p_true    = p_true,
      cp_lower  = cp_lower,
      gamma_fin = gamma_fin
    )
  })
```

```{r cp-post-final-table}
#| label: tbl-cp-post-final
#| tbl-cap: "Symmetric comparison at the closest-to-$\\alpha$ $\\text{CP}_L$ from §4.4 (since no $\\text{CP}_L$ on the grid controls T1E ≤ $\\alpha$ with the z-test final, this is the natural comparator). Same CP-driven SSR machinery, but the final analysis is swapped from z-test to Bayesian posterior. The posterior final produces a monotone, stepwise T1E curve in $\\gamma_\\text{final}$ (consecutive grid values may share an integer critical count and therefore identical T1E), and exact T1E ≤ $\\alpha$ is achievable across multiple $\\gamma_\\text{final}$ values, a far less irregular calibration surface than the z-test final design (Section 4.1)."

cp_post_t1e <- cp_post_sweep |> filter(p_true == p0) |> select(gamma_fin, t1e = power)
cp_post_pwr <- cp_post_sweep |> filter(p_true == p1) |> select(gamma_fin, pwr = power)
cp_post_join <- cp_post_t1e |> inner_join(cp_post_pwr, by = "gamma_fin") |>
  arrange(gamma_fin)

cp_post_join |>
  transmute(
    `gamma_final` = sprintf("%.3f", gamma_fin),
    `T1E at p0`   = percent(t1e, accuracy = 0.1),
    `Power at p1` = percent(pwr, accuracy = 0.1),
    `T1E <= alpha?` = ifelse(t1e <= alpha, "Yes", "No")
  ) |>
  gt() |>
  cols_label(
    `gamma_final`  = md("$\\gamma_\\text{final}$"),
    `T1E at p0`    = md("T1E at $p_0$"),
    `Power at p1`  = md("Power at $p_1$"),
    `T1E <= alpha?` = md("T1E $\\leq \\alpha$?")
  ) |>
  tab_source_note(md(glue(
    "Comparator: $\\text{{CP}}_L = {cp_pl_use}$ (closest-to-α $\\text{{CP}}_L$ ",
    "from §4.4, used because no $\\text{{CP}}_L$ controls T1E with the $z$-test ",
    "final). Exact enumeration."
  ))) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

```{r four-way-summary}
#| label: tbl-four-way
#| tbl-cap: "Three-of-four-way summary. Two SSR rules (CP-driven vs PP-driven) crossed with two final-analysis rules (z-test vs posterior). Three cells are computed; the fourth (PP-driven SSR with z-test final) is left as future work and is not pictured. The three implemented cells show that swapping the final analysis from z-test to posterior renders both SSR variants calibratable, establishing the final-analysis rule as a sufficient explanation of the original CP design's inflation in this setting, but without separating it from a possible SSR-logic contribution."

# Recover the original CP and Bayesian results for the comparison.
# CP + z-test final: at CP_L = cp_pl_use (closest-to-α since no CP_L is calibrated)
cp_z_t1e <- cp_sweep |>
  filter(p_true == p0, round(cp_lower, 4) == round(cp_pl_use, 4)) |> pull(power)
cp_z_pwr <- cp_sweep |>
  filter(p_true == p1, round(cp_lower, 4) == round(cp_pl_use, 4)) |> pull(power)

# Bayesian + posterior final (original Bayesian): at calibrated gamma_fin_cal
bayes_p_t1e <- bayes_t1e
bayes_p_pwr <- bayes_power

# CP-driven SSR + posterior final: calibrate INDEPENDENTLY (smallest gamma_fin
# in the cp_post_join grid that controls T1E ≤ alpha at p = p0) so the cell
# reflects this design's own naturally-chosen calibration, not the Bayesian's.
cp_post_cal_row <- cp_post_join |>
  filter(t1e <= alpha) |>
  slice_min(order_by = gamma_fin, n = 1)
cp_post_gamma_cal <- if (nrow(cp_post_cal_row) > 0) {
  cp_post_cal_row$gamma_fin
} else {
  max(gamma_fin_grid)  # fallback
}
cp_p_t1e <- cp_post_cal_row$t1e[1]
cp_p_pwr <- cp_post_cal_row$pwr[1]

fmt <- function(t1e, pwr) {
  if (is.na(t1e) || is.na(pwr)) return("—")
  sprintf("%s / %s", percent(t1e, accuracy = 0.1), percent(pwr, accuracy = 0.1))
}

tribble(
  ~`SSR rule`,                     ~`z-test final (alpha = 0.05)`,    ~`Posterior final (gamma_fin)`,
  "CP-driven SSR (zone classifier)",  fmt(cp_z_t1e, cp_z_pwr),           fmt(cp_p_t1e, cp_p_pwr),
  "PP-driven SSR (PP classifier)",    "— (future work)",                  fmt(bayes_p_t1e, bayes_p_pwr)
) |>
  gt() |>
  cols_label(
    `SSR rule`                       = "SSR rule",
    `z-test final (alpha = 0.05)`    = md("$z$-test final ($\\alpha = 0.05$)"),
    `Posterior final (gamma_fin)`    = md("Posterior final ($\\gamma_\\mathrm{final}$)")
  ) |>
  tab_source_note(md(glue(
    "Cells report exact T1E at $p_0$ / exact power at $p_1$. ",
    "CP-driven + $z$-test: $\\text{{CP}}_L = {cp_pl_use}$ (closest-to-α from §4.4); ",
    "CP-driven + posterior: $\\gamma_\\text{{final}} = {sprintf('%.3f', cp_post_gamma_cal)}$ ",
    "(this design's own calibrated $\\gamma_\\text{{final}}$); ",
    "PP-driven + posterior: $\\gamma_\\text{{final}} = {sprintf('%.3f', gamma_fin_cal)}$ ",
    "(from §4.4)."
  ))) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(100), table.font.size = "small",
              heading.title.font.size = "small",
              heading.subtitle.font.size = "footnotesize")
```

The implemented comparison establishes a *narrower* claim than a
fully-symmetric two-by-two decomposition would: switching the final
analysis from a $z$-test to a Bayesian posterior is *sufficient* to
restore calibratable T1E in the CP-driven SSR design, even with the
same CP-zone SSR logic. This means the discrete $z$-test final is
**a sufficient cause** of the original CP design's inflation in this
setting. It does not establish that the $z$-test final is the *only*
cause; the complementary variant (PP-driven SSR with a $z$-test
final) is not implemented here, so we cannot rule out that the
PP-zone SSR logic *also* contributes some T1E control on top of
what the posterior final provides. We expect that variant would
inherit the $z$-test discreteness bias (since its final test matches
the original CP design's), but verifying this is left as future work.

For practitioners, the narrower claim is still actionable:
those who want to keep the familiar CP-promising-zone SSR machinery
can recover calibrated T1E by replacing the $z$-test final with a
Bayesian posterior threshold or with an exactly-calibrated critical
count; alternatively, they can adopt a Bayesian PP design end-to-end.


# Discussion

**Why the Mehta–Pocock guarantee fails for single-arm binary trials.**
The analytic Type I error control of @mehta2011 is inherited from the
inverse-normal combination test of @lehmacher1999 and @muller2001,
which combines pre- and post-interim test statistics on a *continuous*
$z$-statistic scale. The combination weights are fixed before the
trial and do not depend on the SSR decision, which preserves the joint
distribution of the combined statistic under $H_0$. In single-arm
binary trials, the interim "test statistic" is a binomial response
count with $n_\text{int} + 1$ possible values, and the final-analysis
$z$-test is applied to the *integer* total response count. The
discrete critical count $x_\text{crit}(N)$ is the smallest integer
with $z \geq z_\alpha$, so the actual fixed-$N$ rejection probability
under $H_0$ generally differs from the nominal $\alpha$ by a *signed*
amount that depends on $(p_0, N)$, it can be either upward (e.g.,
the base setting at $p_0 = 0.23, N = 84$) or downward (e.g., the
$p_0 = 0.10, N = 55$ setting in @fig-generalizability). The promising-zone SSR
moves trials between such finite-$N$ tests, with their own
opposite-sign discreteness biases, in a way that depends on the
$\text{CP}_L$ threshold and the realised $x_\text{int}$. The
resulting T1E behaviour is design-specific and irregular: it is not
a numerical artefact, but a structural consequence of running an
integer-grid test at finite $N$ combined with $N$-changing SSR, and
it cannot be controlled by a fixed-$\text{CP}_L$ rule that
generalises across designs.

**Practical implication for trial designers.** Practitioners using the
CP promising-zone SSR for single-arm binary trials should not rely on
the analytic Type I error guarantee from @mehta2011. In our base
setting no $\text{CP}_L$ on the grid achieves
$\text{T1E} \leq \alpha$; in other design configurations (@fig-generalizability),
the picture is mixed, some $\text{CP}_L$ values control T1E, others
do not, and there is no consistent rule for which. "Tuning"
$\text{CP}_L$ on a fixed grid is therefore unreliable: the
discreteness bias dominates the SSR effect and changes sign across
designs, so a $\text{CP}_L$ that controls T1E in one configuration
may not control it in another. Practitioners who must nonetheless
use this design should report exact T1E (or simulated T1E with
adequate precision) across their full $\text{CP}_L$ grid for their
*specific* design parameters, treat the
worst-case T1E across plausible design perturbations as the operating
characteristic, and not extrapolate calibration from one design to
another.

**Bayesian PP alternative.** Where calibrated Type I error control
is required, the Bayesian PP SSR design with separate
$\gamma_\text{int}$ and $\gamma_\text{final}$ thresholds offers an
attractive alternative. Exact Type I error is monotone in
$\gamma_\text{final}$, and although it remains stepwise (success
ultimately depends on integer counts), the calibration surface is
far less irregular than the $\text{CP}_L$ surface, enumeration
over the $\gamma_\text{final}$ grid yields a clean smallest-value
that controls T1E, with no need to navigate sign-changing
discreteness bias. Power and operating characteristics are
competitive with the best-calibrated CP design (when one exists)
and dominate it when CP cannot be calibrated.

**Regulatory implication.** This is consistent with the FDA 2019 adaptive
design guidance [@fda2019], which requires demonstration of Type I
error control, analytically where possible, otherwise via simulation.
For the one-look single-arm binary design we study, exact enumeration
is in fact available (we use it for every headline number above) and
provides stronger evidence than simulation can; for more complex
adaptive designs where enumeration becomes impractical (multiple
interim looks, continuous endpoints with adaptive boundaries),
simulation remains the standard route. Our results imply that any
analysis, exact or simulated, must evaluate the entire $\text{CP}_L$
grid the design might use, not a single point.

**Limitations.** The designs studied here have a single interim look,
a binary ORR endpoint, a Jeffreys non-informative prior, and a
one-sided normal-approximation $z$-test for the final analysis; our
exact enumeration is specific to that combination. The qualitative
claim, that the discrete $z$-test critical count produces a
*design-specific* (and signed) rounding bias at finite $N$, and that
$N$-changing SSR rearranges these biases in a way that defies
fixed-rule calibration, should generalise to other discrete
endpoints where the final-analysis test reduces to integer-grid
arithmetic. An exact (Clopper–Pearson) final analysis would push T1E
uniformly below $\alpha$ at the cost of obscuring the SSR-driven
boundary-crossing effects we characterise; the choice between
approximate and exact tests is a separate design decision that
practitioners should make explicitly. Multiple interim looks,
informative priors (e.g., MAP), and a fully-symmetric
Bayesian-vs-frequentist comparison (in which the two designs share a
final analysis rule and differ only in their interim decision logic)
are natural directions for follow-up work; for multiple-look designs
the simulator-based `run_*_scenario` framework in this manuscript is
the natural extension path, since exact enumeration becomes
combinatorially expensive once the look count grows.


# Conclusion

The Mehta–Pocock conditional-power promising-zone SSR with a one-sided
$z$-test final analysis cannot be calibrated to a target Type I error
in single-arm binary trials by a rule that generalises across designs.
In a representative oncology setting ($p_0 = 0.23$, $p_1 = 0.35$,
$N_\text{init} = 84$, $n_\text{int} = 65$, $N_\text{max} = 200$), exact
Type I error sits above the nominal $\alpha$ at every $\text{CP}_L$
on the grid; across other design configurations, the discrete
$z$-test critical count's rounding bias changes sign, so
$\text{CP}_L$ values that control T1E in one configuration may not
control it in another. The overall T1E surface is too irregular to
support a fixed-rule calibration. The partial symmetric comparison
in Section 4.6 shows that switching the CP design's final analysis
to a Bayesian posterior threshold is **sufficient** to restore
calibratable T1E, even with the same CP-zone SSR logic; the
discrete $z$-test final is therefore a *sufficient cause* of the
inflation. Because we do not implement the complementary variant
(PP-driven SSR with a $z$-test final), we cannot rule out an
additional contribution from the SSR-logic itself. Practitioners
who want calibrated Type I error control therefore have two
practical paths: replace the $z$-test final with a calibrated
final-analysis rule (Bayesian posterior threshold or an
exactly-calibrated critical count chosen so that the fixed-$N$
test attains nominal $\alpha$) while keeping the CP-zone SSR
machinery, or adopt a Bayesian PP SSR design end-to-end. Bayesian
PP SSR is a particularly attractive alternative because its exact
Type I error is monotone in $\gamma_\text{final}$ and calibrates
directly by enumeration, the curve is still stepwise (success
ultimately depends on integer response counts), but the
calibration surface is far less irregular than the CP_L surface.


# Code and Data Availability

All code used in this paper is released as a public R/Quarto repository,
presented anonymously for double-blind peer review at
<https://anonymous.4open.science/r/CP-SSR-Binary-Trials-8DB8/>. The
non-anonymized public GitHub repository will be disclosed in the
camera-ready version after acceptance. The repository contains the Quarto
manuscript source (`CP_PromisingZone_SSR_TypeIError.qmd`),
the bibliography (`references.bib`), and an R script (`R/analysis.R`)
that defines the exact-enumeration helpers (`cp_oc_exact`, `bayes_oc_exact`,
`cp_oc_post_final`, `fixed_n_t1e`) and simulator-based fallback functions
(`sim_trial_cp`, `sim_trial_bayes`, `run_*_scenario`). The manuscript is
fully reproducible from the `.qmd` source with R $\geq$ 4.4 and the
following CRAN packages: `tidyverse`, `gt` ($\geq$ 1.0.0), `scales`,
`patchwork`, `glue`, and `latex2exp`. No external data are used: every
numerical result is computed from the exact enumeration code in
`R/analysis.R`.


# References

:::{#refs}
:::


<!--
Session info is omitted from the journal-submission PDF. The exact R
version and package versions used to render this manuscript are pinned
in the public repository's `R/analysis.R` and reproducible via the
included `.qmd` source. To re-enable session info during development,
delete the surrounding HTML comment markers.

```{r session}
#| eval: false
sessionInfo()
```
-->
